---
title: "Trial"
output: html_notebook
---


```{r}
suppressPackageStartupMessages({
library(MASS)})
```

We create functions for doing mds and getting labels the traditional way given the centred data and genes

```{r}
do.mds <- function(df, genes, labels, title)
{
  d <- dist(t(df[genes,]) )
  mds.single.cell <- isoMDS(d)
  mds.single.cell <- data.frame(mds.single.cell)
  colnames(mds.single.cell) = c('dim1', 'dim2')
  mds.single.cell$lab <- labels
  return(ggplot(data = mds.single.cell, aes_string(x = 'dim1', y = 'dim2',
                                            color = 'lab' ))+
    geom_point(size=3)+ggtitle(title)  )
}

get.understanding <- function(centered.data, mds.data, genes, title)
{
  cov.df <- get.covariance(centered.data$single.cell, centered.data$bulk, 
                           genes)
  labels <- get.labels(cov.df)
  table(labels)
  plot <- do.mds(mds.data, genes, labels, title)
  return(list(cov.df = cov.df, labels = labels, plot = plot))
}

```

First we try to see the MDS plots with different thresholds of slope. We also try to see what happens if we take the entire gene set.

Function for getting genes for threshold
```{r}
get.marker.genes <- function(df.spread.diff, threshold, genes.common)
{
  g.o.i <- sapply(df.spread.diff, 
                function(testis.type) {
                  get.points.beyond(testis.type, threshold)
                },
                simplify=FALSE )
#g.o.i <- unname(g.o.i)
return(sapply(g.o.i, function(g) genes.common[g], simplify = F))
}
```

Getting genes for different thresholds
```{r}
g.2 <- get.marker.genes(testis.type.spread.diff, 2, genes.common)
g.1.5 <- get.marker.genes(testis.type.spread.diff, 1.5, genes.common)
g.1 <- get.marker.genes(testis.type.spread.diff, 1, genes.common)
```

Multiplot function
```{r}
multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  
  # Make a list from the ... arguments and plotlist
  if(length(...) == 1)
  {
    print('aye')
    plots <- c(list(...), plotlist)
  }
  else
    plots <- c(..., plotlist)
  #print(plots)
  numPlots = length(plots)
  
  # If layout is NULL, then use 'cols' to determine layout
  if (is.null(layout)) {
    # Make the panel
    # ncol: Number of columns of plots
    # nrow: Number of rows needed, calculated from # of cols
    layout <- matrix(seq(1, cols * ceiling(numPlots/cols)),
                     ncol = cols, nrow = ceiling(numPlots/cols))
  }
  
  if (numPlots==1) {
    print(plots[[1]])
    
  } else {
    # Set up the page
    grid.newpage()
    pushViewport(viewport(layout = grid.layout(nrow(layout), ncol(layout))))
    
    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      # Get the i,j matrix positions of the regions that contain this subplot
      matchidx <- as.data.frame(which(layout == i, arr.ind = TRUE))
      
      print(plots[[i]], vp = viewport(layout.pos.row = matchidx$row,
                                      layout.pos.col = matchidx$col))
    }
  }
}

```

Below we find labels and the plots for our respective genes
```{r}
p.all.genes <- get.understanding(centered.data,log2(testis.single.cell.gmpr+1),
                                 genes.common, 'all genes')
p.2.genes <- get.understanding(centered.data,log2(testis.single.cell.gmpr+1),
                              unlist(g.2), 'Threshold 2')
p.1_5.genes <- get.understanding(centered.data,log2(testis.single.cell.gmpr+1),
                               unlist(g.1.5), 'Threshold 1.5')
p.1.genes <- get.understanding(centered.data,log2(testis.single.cell.gmpr+1),
                               unlist(g.1), 'Threshold 1')

multiplot(p.all.genes$plot, p.2.genes$plot, p.1_5.genes$plot, p.1.genes$plot,
          cols = 2)

sapply(list(p.all.genes, p.2.genes, p.1_5.genes, p.1.genes), function(p)
  {
  table(p$labels)
  })

```
The thing worth noting is that from threshold 2 to 1 number of sd rise and sz fall but both still are the highest which confirms with the paper. However sl also increases which is not good.


Another thing worth seeing was looking at overall count of the marker genes in single cell.
```{r}
#Genes with reasonable count in each cell
g.reas <- apply(testis.single.cell.gmpr, 1, function(r)
{
  sum(r >= 10)
})
g.reas.length <- which(g.reas >= 40)

#Genes with rowMean greater than 5
g.mean <- which(rowMeans(testis.single.cell.gmpr) >= 10)

g.high <- genes.common[intersect(g.reas.length, g.mean)]
```

Using the above and intersecting with our marker set
```{r fig.height=30, fig.width=10}
marker.high <- sapply(list(g.1, g.1.5, g.2), function(g) sapply(g, function(g.sub) intersect(g.sub, g.high), simplify = F), simplify = F)
names(marker.high) <- paste('marker', c('1', '1.5', '2'), sep = '')
p.marker.high <- sapply(seq_along(marker.high), function(i)
  {
  get.understanding(centered.data,log2(testis.single.cell.gmpr+1),
                                 unlist(marker.high[[i]]), names(marker.high)[i])
  }, simplify = F)
multiplot(p.marker.high[[1]]$plot, p.marker.high[[2]]$plot, p.marker.high[[3]]$plot, col =1 )
sapply(p.marker.high, function(p.high) table(p.high$labels))

```


Having a look at Florent's Markers
```{r}
##Loading the data
single.cell.markers <- read.table('~/Dropbox/intern/data/markers_single_cell.txt', header = T)
head(single.cell.markers)
```

```{r}
#Finding labels and doing an MDS plot using Florent's Markers
p.flo <- get.understanding(centered.data, log2(testis.single.cell.gmpr+1),
                      intersect(single.cell.markers$ENSEMBL, genes.common), 'Flo Markers')

#Doing an MDS using Florent's labels
p.flo.labs <- do.mds(log2(testis.single.cell.gmpr+1), 
       intersect(single.cell.markers$ENSEMBL, genes.common),
       labels.single.cell$rna,
       title = 'Flo Labels')

multiplot(p.flo$plot, p.flo.labs)
```

Having a look at Florent's markers in our diff vs spread plot
```{r}
##Function for plotting given genes across the diff vs spread plots for all the types
get.look <- function(df.spread.diff, type, threshold, markers)
{
  par(mfrow = c(3, 2))
  for(i in seq_along(df.spread.diff))
  {
    plot(df.spread.diff[[i]], 
         main = names(df.spread.diff)[i])
    abline(c(0,threshold))
    abline(c(0,-threshold))
    abline(h = c(-2,2))
    inds = match(markers, 
               rownames(df.spread.diff[[i]]))
    points(df.spread.diff[[i]]$spread[inds],
           df.spread.diff[[i]]$diff[inds],
           col = 'red')
  }
  title(type, outer = T)  
}
```

Plots
```{r}
get.look(testis.type.spread.diff, 'SZ', 2, single.cell.markers$ENSEMBL[1:10] )
get.look(testis.type.spread.diff, 'SD', 2, single.cell.markers$ENSEMBL[11:30] )
get.look(testis.type.spread.diff, 'SC', 2, single.cell.markers$ENSEMBL[31:40] )
get.look(testis.type.spread.diff, 'SL', 2, single.cell.markers$ENSEMBL[41:50] )
get.look(testis.type.spread.diff, 'SG', 2, single.cell.markers$ENSEMBL[51:60] )
```

Trying correlation as a metric to get labels
```{r}
get.cor.cov <- function(single.cell.data, bulk.data, genes)
{
  #print(genes)
  return(cor(single.cell.data[genes, ], bulk.data[genes, ], method = 'spearman'))
}

cor.df <- get.cor.cov(centered.data$single.cell, centered.data$bulk, unlist(g.2))
cor.df <- get.cor.cov(log2(testis.single.cell.gmpr+1), testis.types.rna.log, unlist(g.2))
cor.labels <- get.labels(cor.df)
table(cor.labels)
```

```{r}
##Using only marker genes
get.cov.using.marker.genes <- function(single.cell, bulk.data, g.o.i)
{
  sapply( seq_along(g.o.i), function(i)
   (bulk.data[ g.o.i[[i]], i ] %*% single.cell[ g.o.i[[i]], ] )/length(g.o.i[[i]]), USE.NAMES = T)

}
cov.2.marker <- get.cov.using.marker.genes(centered.data$single.cell, centered.data$bulk, g.2)
colnames(cov.2.marker) = names(g.2)
cov.2.labels <- get.labels(cov.2.marker)
table(cov.2.labels)
sum(cov.2.labels == labels)
do.mds(log2(testis.single.cell.gmpr+1), unlist(g.2), cov.2.labels, 'A')
```


Plotting covariance of each cell type vs each other
```{r}
p.list <- list()
#par(mfrow = c(2,5))
k = 1
for(i in seq(ncol(covariances)))
{
  if(i == ncol(covariances))
    break()

  for(j in c((i+1):ncol(covariances)))
  {
    df = data.frame(col1 = p.marker.1$cov.df[,i], col2 =  p.marker.1$cov.df[,j],
                    labs = labels.rank)
    p <- ggplot(data = df, aes_string(x = 'col1', y = 'col2',
                          color = 'labs' ))
      p <- p + geom_point(size=1)
      p <- p + ggtitle(paste(colnames(p.marker.1$cov.df)[i],
                                              colnames(p.marker.1$cov.df)[j], sep='_'))
      p <- p + coord_fixed()
      
    p.list[[k]] <- p         
    k = k + 1
    
  }
}

multiplot(p.list, cols = 5)

```

Looking at high marker genes
```{r}
g.reas <- apply(testis.single.cell.gmpr, 1, function(r)
{
  sum(r >= 100)
})
g.reas.length <- which(g.reas >= 50)

g.marker.1 <- intersect(unlist(g.1), genes.common[g.reas.length])
p.marker.1 <- get.understanding(centered.data, log2(testis.single.cell.gmpr+1),
                                  g.marker.1, 'Marker 1')
p.marker.1$plot
length(g.marker.1)
table(p.marker.1$labels)
#Genes with rowMean greater than 5
g.mean <- which(rowMeans(testis.single.cell.gmpr) >= 10)

g.high <- genes.common[intersect(g.reas.length, g.mean)]
```

Using rank based approach
```{r}
library(geneplotter)
multidensity(p.marker.1$cov.df)
df.marker.rank.1 <- apply( p.marker.1$cov.df, 2, rank )
labels.rank <- get.labels(df.marker.rank.1)
table(labels.rank)
```

Dropouts
```{r}
dropout <- apply(testis.single.cell, 1, function(r)
  {
  mean(r == 0)
})
sum(dropout < 0.7)

anna <- read.csv(file = 'data/GSE67833_Gene_expression_matrix.csv', row.names = 1)
drop.anna <- apply(anna, 1, function(r)
  {
  mean(r == 0)
})
sum(drop.anna < 0.7)
par(mfrow = c(1,2))
hist(dropout)
hist(drop.anna,
     ylim = c(0,8000))
multiecdf(list(drop.anna, dropout))

plot( sort(drop.anna), seq_along(drop.anna), type="l", ylim=c(0,20000) )
lines( sort(dropout), seq_along(dropout), type="l", col="red" )
abline(h=(0:4)*10000,col="#00000060")

sum(names(dropout) == genes.common) == length(genes.common)
abline(v = dropout[match(intersect(single.cell.markers$ENSEMBL, genes.common),
      genes.common)], col="#0000FF30")

sum(labels.single.cell$V1 == colnames(testis.single.cell)) == ncol(testis.single.cell)
df.drop.samples <- data.frame(drop.samples = apply(testis.single.cell, 2, function(col) mean(col == 0 )),
  cluster = labels.single.cell$rna, row.names = colnames(testis.single.cell))
boxplot(data = df.drop.samples, drop.samples~cluster)

```